MWA tied-array processing III: Microsecond time resolution via a polyphase synthesis filter

### 2020-05-08 ipfb.tex
Updated draft of paper text in partial response of the referee's comments

@@

\subsection{Design}

The spectral output of the PFB, $X_k[m]$, can be used to reconstruct the time series $\hat{x}[n]$ by means of a \textit{synthesis filter}.
The hat signifies the fact that the resulting time series does not necessarily equal the original input to the PFB, $x[n]$, since one is free to choose the synthesis filter itself, $f$:
\begin{equation}
    \hat{x}[n] = \sum_{m = 0}^{M-1} f[n - mM]\,
        \frac{1}{K} \sum_{k=0}^{K-1} X_k[m]\,e^{2\pi jkn/K}.
    \label{eqn:ipfb}
\end{equation}
The necessary and sufficient condition for exact equality, $\hat{x}[n] = x[n]$, for a critically sampled PFB is \citep[proven in][]{Crochiere1983}:
\begin{equation}
    \lambda(s) = \sum_{m=0}^{M-1} f[n-mM]\,h[mM - n + sM] = \Kron
    \label{eqn:invertibility_condition}
\end{equation}
for all values of $n$, where $\Kron$ is the Kronecker delta.
Eq. \eqref{eqn:invertibility_condition} formally resembles a kind of ``convolution'' of the synthesis filter with the analysis filter, where the translational part of the convolution operation is restricted to step sizes of $M$.

@@

For the analysis filter $h[n]$ defined in Eq. \eqref{eqn:filter}, it is found that choosing the synthesis filter
\begin{equation}
    f[n] = \frac{h[-n]}{H}
    \label{eqn:synthesisfilter}
\end{equation}
where $H = \sum_{n=0}^{N-1} h[n]$ is a normalising factor, meets the invertibility condition sufficiently closely for our purposes, as demonstrated below.
With the analysis and synthesis filters thus defined, the value of Eq. \eqref{eqn:invertibility_condition} can be calculated and is illustrated in Fig. \ref{fig:inverse_condition}.
It is readily seen that $\lambda(s)$ closely approximates, but does not exactly equal, $\Kron$.
The quantitative effect of this analysis-synthesis pair is discussed in Section \ref{sec:fidelity}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{inverse_condition.eps}
    \caption{The condition for invertibility, Eq. \eqref{eqn:invertibility_condition}, demonstrated for the analysis filter given in Eq. \eqref{eqn:filter} and the synthesis filter given in Eq. \eqref{eqn:synthesisfilter}. The fact that $\lambda(s)$ only approximates the delta function gives rise to spurious artefacts (``temporal echoes'') in the time domain, which can be seen in Fig. \ref{fig:impulse_response}.}
    \label{fig:inverse_condition}
\end{figure}

>>>>

\newcommand{\changes}[1]{{\color{red}#1}}

@@

In order to regain the time resolution lost during analysis, the spectral output of the PFB must be transformed back into the time domain.
The inverse to the analysis operation defined in Eq. \eqref{eqn:wdft} converts a spectrum $X_k[m]$ into a time series $\hat{x}[n]$ by means of a \textit{synthesis filter}, $f[n]$:
\begin{equation}
    \hat{x}[n] = \sum_{m = -\infty}^{\infty} f[n - mM]\,
        \frac{1}{K} \sum_{k=0}^{K-1} X_k[m]\,e^{2\pi jkn/K}.
    \label{eqn:ipfb}
\end{equation}
In this expression, the index $m$ is allowed to run over all integers in order accommodate an arbitrarily large synthesis filter.

Back-to-back analysis-synthesis filters can be designed so that the original time series can be perfectly reconstructed without any loss of signal (i.e. $\hat{x}[n] = x[n]$ exactly), and the system as a whole can be thought of as an identity operation.
The condition for perfect reconstruction can be found by substituting the analysed spectrum obtained from Eq. \eqref{eqn:wdft} into the synthesis operation defined in Eq. \eqref{eqn:ipfb}.
It can be shown that, for a critically sampled system,
\begin{equation}
    \begin{aligned}
        \hat{x}[n]
            &= \sum_{s = -\infty}^{\infty} x[n - sM]
               \sum_{m = -\infty}^{\infty} f_m[n]\, h_m[sM - n] \\
            &= x[n] \sum_{m = -\infty}^{\infty} f_m[n]\, h_m[-n] +{} \\
            &\quad \sum_{\substack{s = -\infty \\ s \ne 0}}^{\infty} x[n - sM]
               \sum_{m = -\infty}^{\infty} f_m[n]\, h_m[sM - n]
    \end{aligned}
    \label{eqn:backtoback}
\end{equation}
where, following \citet{Crochiere1983}, we have adopted the shorthand notation
\begin{equation}
    \begin{aligned}
        h_m[n] &\equiv h[n - mM], \\
        f_m[n] &\equiv f[n - mM].
    \end{aligned}
\end{equation}
It can be seen that perfect reconstruction corresponds to the case when the only non-zero contribution comes from the $s = 0$ term, giving rise to the necessary and sufficient condition
\begin{equation}
    \sum_{m = -\infty}^{\infty} f[n-mM]\,h[(m - s)M - n] = \Kron
    \label{eqn:invertibility_condition}
\end{equation}
for all values of $n$, where $\Kron$ is the Kronecker delta.

Finding the synthesis filter that perfectly inverts a given analysis filter is tantamount to solving Eq. \eqref{eqn:invertibility_condition} for $f[n]$.
However, exact solutions do not always exist, and in general, numerical methods must be employed to find a synthesis filter that minimises the reconstruction error.

\subsection{Reconstruction error}

In the context of an astrophysical signal, it is desirable to quantify the reconstruction error in terms of the signal-to-noise ratio (S/N).
If we assume that individual samples follow, and are dominated by, the same Gaussian noise statistics, then the reduction in S/N of a reconstructed sample can be estimated by considering the fraction of the power of the reconstructed sample that came from the original sample:
\begin{equation}
    \frac{[\text{S/N}]_\text{recon}}{[\text{S/N}]_\text{orig}}
      = \frac{\left(\sum\limits_{m = -\infty}^{\infty} f_m[n]\,h_m[-n]\right)^2}{
            \sum\limits_{s = -\infty}^{\infty}
            \left(\sum\limits_{m = -\infty}^{\infty} f_m[n]\,h_m[sM - n]\right)^2
        }
\end{equation}
This expression is invariant under the substitution $n \rightarrow n + \lambda M$, $\lambda \in \mathbb{Z}$, which indicates that the (average) reconstruction error is only a function of where the sample in question falls within a tap.
Consequently, any synthesis filter (except the exact inverse of the analysis filter, if it exists) will introduce a ``ringing'' effect into the reconstructed time series that has a period equal to the size of the PFB tap.

\subsection{Design constraints}

Eq. \eqref{eqn:backtoback} suggests that a reconstructed time sample potentially contains contributions from up to $P$ original samples, spaced at intervals of one tap.

@@

For the analysis filter $h[n]$ defined in Eq. \eqref{eqn:filter}, it is found that choosing the synthesis filter
\begin{equation}
    f[n] = \frac{h[-n]}{H}
    \label{eqn:synthesisfilter}
\end{equation}
where $H = \sum_{n=0}^{N-1} h[n]$ is a normalising factor, meets the invertibility condition sufficiently closely for our purposes, as demonstrated below.
With the analysis and synthesis filters thus defined, the value of Eq. \eqref{eqn:invertibility_condition} can be calculated and is illustrated in Fig. \ref{fig:inverse_condition}.
It is readily seen that $\lambda(s)$ closely approximates, but does not exactly equal, $\Kron$.
The quantitative effect of this analysis-synthesis pair is discussed in Section \ref{sec:fidelity}.

%\begin{figure}
%    \centering
%    \includegraphics[width=\columnwidth]{inverse_condition.eps}
%    \caption{The condition for invertibility, Eq. \eqref{eqn:invertibility_condition}, demonstrated for the analysis filter given in Eq. \eqref{eqn:filter} and the synthesis filter given in Eq. \eqref{eqn:synthesisfilter}. The fact that $\lambda(s)$ only approximates the delta function gives rise to spurious artefacts (``temporal echoes'') in the time domain, which can be seen in Fig. \ref{fig:impulse_response}.}
%    \label{fig:inverse_condition}
%\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{snr.eps}
    \caption{The reconstructed S/N as a function of the position of the reconstructed sample within a tap.}
    \label{fig:snr}
\end{figure}
